\paperTitle

### 2014-01-31 doc/icml2014/intro.tex
Added some figures and made some changes to sample complexity stuff
Percy's changes

% 1. Latent variable models are good.
Latent variable graphical models offer a succinct representation of
a rich model family. 
% 2. Learning them is hard.
Despite their success across many fields
\cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype,liang06discrimative,kirkpatrick10painless},
learning these models remains a difficult problem due to the
non-convexity of the likelihood. Local optimization (e.g.
expectation-maximization) is the standard approach, but is susceptible
to local optima.

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the method of moments and
spectral decomposition have offered a refreshing and promising perspective on
this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
These methods exploit the linear algebraic properties of the model and
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods apply to a limited set of models, and are thus
not as universally applicable as EM.

\begin{figure}[t]
  \label{fig:approach}
  \centering
  \subimport{figures/}{approach.tikz}
  \caption{Overview of approach}
\end{figure}

% 4. State what we do: exploit moment constraints to make the problem easier.
In this work, we exploit the spectral method to learn moment constraints
  on the observed variables and show how these constraints lead to consistent parameter estimates 
  for any graphical model with the property that there exist at least
  three ``views'' or conditionally independent observed variables for
  every hidden variable (\propertyref{bottleneck}).

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
Our approach is illustrated in \figureref{approach}. 
First, we identify the views $x_1, x_2, x_3$ for each hidden variable $h$. 
We then use the tensor factorization algorithm of
  \citet{anandkumar12moments,anandkumar13tensor} to identify the
  conditional moments.
Given these factors, we show that there exists a set of observed
  variables for every clique such that the likelihood is convex in the
  parameters of that clique. 
This guarantees that we can recover the true marginal distributions
  using EM; finally, with some post-processing, we are able to
  consistently recover the true parameters of the model.
In \sectionref{directed}, we detail our algorithm and describe the
  settings and technical assumptions in which this approach is guaranteed
  to work.
As a particular example, we are able to recover parameters for the
  grid model depicted in \figureref{approach} under the condition that
  $O$ and $T$ is full-rank.

Our approach also applies to undirected log-linear latent variable
  models, where we use the recovered marginals to make the learning
  problem convex (\sectionref{undirected}).
Finally, our approach gracefully extends to more general model families
  in which we are unable to recover the marginal distributions for all
  cliques; in such scenarios, we use posterior regularization to constrain
  EM.

>>>>

% 1. Latent variable models are good.
Latent-variable graphical models provide compact representations of data
and have been employed across many fields [TODO: find a much better/broader set of citations]
\cite{quattoni04crf,beaumont2004bayesian,haghighi06prototype}.
% 2. Learning them is hard.
However, learning these models remains a difficult problem due to the
non-convexity of the negative log-likelihood.  Local methods such as Expectation Maximization (EM)
are the norm, but are susceptible to local optima.

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the spectral method of moments
have offered a refreshing and promising perspective on
this learning problem
\citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
[TODO: make sure we are citing the right people]
These methods exploit the linear algebraic properties of the model to
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods apply to a limited set of models, and are thus
not as broadly applicable as EM.

\begin{figure}[t]
  \label{fig:approach}
  \centering
  \subimport{figures/}{approach.tikz}
  \caption{[TODO: need three steps: tensor factorization, optimize composite likelihood, recover model parameters; Overview of approach}
\end{figure}

% DONE: need to say from the beginning that we have a hybrid approach, not just using method of moments;
% there are really two contributions: one is the consistent estimate result for class of models,
% and the other is the general idea of using partial constraints]

% 4. State what we do: exploit moment constraints to make the problem easier.
In this paper,
we show that a much broader class of directed and undirected graphical models can be consistently estimated,
specifically those in which each latent variable has three conditionally
independent observed variables (``views'')---see \propertyref{bottleneck}.
The key idea is to leverage the spectral method of moments,
not to directly provide a consistent parameter estimate as in previous work,
but as partial constraints on a likelihood-based objective.
%For a broad class of directed and undirected graphical models

%In this work, we exploit the spectral method to learn moment constraints
  %on the observed variables and show how these constraints lead to consistent parameter estimates 
  %for any graphical model with the property that there exist at least
  %three ``views'' or conditionally independent observed variables for
  %every hidden variable (\propertyref{bottleneck}).

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
Our approach is illustrated in \figureref{approach}. 
First, we identify the views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar12moments,anandkumar13tensor} to identify the \emph{conditional
moments} $\E[x_i^a \mid h_i]$ and $\E[x_i^b \mid h_i]$ for each $i$.
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
each latent clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM.
Finally, the latent marginals yield the model parameters
via simple post-processing in the directed case and convex optimization in the undirected case.
%Given these factors, we show that there exists a set of observed
  %variables for every clique such that the likelihood is convex in the
  %parameters of that clique. 
%This guarantees that we can recover the true marginal distributions
  %using EM.
  %finally, with some post-processing, we are able to
  %consistently recover the true parameters of the model.
In \sectionref{directed}, we detail our algorithm and
show that our algorithm yields consistent parameter estimates
under non-degenerate rank assumptions on the conditional moments.
%describe the
  %settings and technical assumptions in which this approach is guaranteed
  %to work.
% PL: hard to talk about $O$ and $T$ at this stage.
%As a particular example, we are able to recover parameters for the
  %model depicted in \figureref{approach} if $O$ and $T$ are full-rank.

%Our approach also applies to undirected log-linear latent variable
  %models, where we use the recovered marginals to make the learning
  %problem convex (\sectionref{undirected}).
Finally, our approach gracefully extends to more general model families
  in which we are unable to recover the marginal distributions for all
  hidden cliques; in such scenarios, we use posterior regularization to constrain
  EM.  We show that empirically, using such constraints alleviates local
  optima issues.

### 2014-01-31 doc/icml2014/related.tex
Percy's changes

Parameter estimation for latent variable models using the method of
  moments has received much attention in last few years.
Our work uses the \TensorFactorize algorithm presented by
  \citet{anandkumar13tensor}, though \citet{anandkumar12moments} first
  studied the consistent parameter estimation algorithm for the
  three-view mixture model.
%\todo{What about Mossel and Roch?}
Several authors have extended this work to settings that do not meet the
  criterion \propertyref{bottleneck}, but rely on knowing more about
  the underlying correlation structure of the problem.
\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar12linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered. 
In comparison, our work applies to a general family of models, including
  latent tree models, directed and undirected grids, with no assumptions
  on the correlation structure aside from independences encoded in the
  graph.

>>>>

Parameter estimation for latent variable models using the method of
  moments has received much attention in last few years.
Our work uses the \TensorFactorize algorithm presented by
  \citet{anandkumar13tensor}, though \citet{anandkumar12moments} first
  studied the consistent parameter estimation algorithm for the
  three-view mixture model.
%\todo{What about Mossel and Roch?}
Several authors have extended this work to settings that do not meet the
  criterion \propertyref{bottleneck}, but rely on knowing more about
  the underlying correlation structure of the problem.
\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar2013linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered. 
In comparison, our work applies to a general family of models, including
  latent tree models, directed and undirected grids, with no assumptions
  on the correlation structure aside from independences encoded in the
  graph.

### 2014-01-31 doc/icml2014/undirected.tex
Updated figures and some details and notation
Percy's changes

As before, let $\sG$ be a graphical model with observed variables $X$ and hidden variables $H$.
We assume that the undirected graphical model is a log-linear latent
  variable model with parameters $\theta$,
\begin{align*}
  p_\theta(\vx, \vh) &= \exp( \sum_\sC \theta_\sC^T \phi_\sC(\vx,\vh) - A(\theta) ),
\end{align*}
where $\phi_\sC(\vx,\vh)$ is a feature vector for the clique $\sC$ and
  $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta_\sC^T \phi_\sC(\vx,\vh) )$ is the
  log-partition function.
While $\phi_\sC(x,h)$ can be an arbitrary featurization, we assume
  without loss of generality that $\phi_\sC(x,h)$ are the indicator
  features $\BI[x,h]$\footnote{Note that $\BI[x,h]$ are sufficient
  statistics for the model, and hence we do not lose any statistical
  efficiency in using them.}.

As before, our objective is to find a set of parameters $\hat \theta$
  that maximize the likelihood of data $\sD$.
\begin{align}
  \hat \theta 
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
\end{align}
Note that this problem is not concave in $\theta$. 
However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
  problem simplifies to,
\begin{align}
  \hat \theta 
  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^T Z_\sC - A(\theta), \label{eqn:obj}
\end{align}
which is concave in $\theta$.
In \sectionref{directed}, we showed that \LearnMarginals is a consistent
  estimator for $Z_\sC$ for any model satisfying \propertyref{exclusive-views}. 
The result is that the output of \algorithmref{undirected}, $\hat\theta$, is a consistent estimator for $\theta^*$
  when the model $p_\theta(\theta)$ is identifiable.

>>>>

As before, let $\sG$ be a graphical model with observed variables $X$ and hidden variables $H$.
We assume that the undirected graphical model is a log-linear latent
  variable model with parameters $\theta$,
\begin{align*}
  p_\theta(\vx, \vh) &= \exp( \sum_\sC \theta_\sC^\top \phi_\sC(\vx,\vh) - A(\theta) ),
\end{align*}
where $\phi_\sC(\vx,\vh)$ is a feature vector for the clique $\sC$ and
  $A(\theta) = \sum_{\vx, \vh}  \exp( \sum_\sC \theta_\sC^\top \phi_\sC(\vx,\vh) )$ is the
  log-partition function.
While $\phi_\sC(x,h)$ can be an arbitrary featurization, we assume
  without loss of generality that $\phi_\sC(x,h)$ are the indicator
  features $\BI[x,h]$\footnote{Note that $\BI[x,h]$ are sufficient
  statistics for the model, and hence we do not lose any statistical
  efficiency in using them.}.

As before, our objective is to find a set of parameters $\hat \theta$
  that maximize the likelihood of data $\sD$.
\begin{align}
  \hat \theta 
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} p_\theta(\vx,\vh) \nonumber \\
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh \in \sH} \exp( \theta^\top\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
\end{align}
Note that this problem is not concave in $\theta$.

\note{Consider the supervised setting.
As we get infinite amounts of data, we get this simple form.
As MoM is a consistent estimator for marginals, we can optimize this
  likelihood and it is consistent.
}

However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
  problem simplifies to,
\begin{align}
  \hat \theta 
  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
\end{align}
which is concave in $\theta$.
In \sectionref{directed}, we showed that \LearnMarginals is a consistent
  estimator for $Z_\sC$ for any model satisfying \propertyref{exclusive-views}. 
The result is that the output of \algorithmref{undirected}, $\hat\theta$, is a consistent estimator for $\theta^*$
  when the model $p_\theta(\theta)$ is identifiable.
